Recognizing Strong Gravitational Lenses

### 2015-02-17 doc/Milestone/milestone.tex
milestone finished! close issue #3. Also a while back I could close issue #4

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\begin{document}

@@

Unfortunately, for how useful strong gravitational lenses are, they are also
extremely rare. A next generation optical survey like the Large Synoptic Survey
Telescope can expect to find only ten thousand lenses in the whole sky, while
it will find ten billion galaxies. Currently in astronomy there are only order
hundreds of strong gravitational lenses known, mostly discovered by 'eyeball
squads' of graduate students. The small number means that target criteria must
be somewhat broad in order to maintain a relatively high completeness. Using
reasonable target criteria to find strong lenses such as looking only at
massive galaxies still means that nearly ten million objects will need to be
inspected in the next generation in order to find those ten thousand lenses. A
team of ten graduate students could expect to spend about 14 years looking at
these objects. Computer algorithms are not much better: most current machine
learning algorithms are woefully-underpowered for this task, and generally have
poor completeness or poor purity -- and often poor both. Additionally, some
algorithms are better at finding some types of lenses than others -- some
perform well on the brilliant arcs, but poorly on the multiply-imaged objects,
or vice versa. New algorithms need to be developed to find more strong
gravitational lenses, and more strong lenses need to be found to power these
algorithms.

{\sc Space\,Warps}\xspace is a citizen-science initiative designed to overcome
these two problems. The program has users examine images from the
Canada-France-Hawaii Telescope Legacy Survey (CFHTLS) and vote on where they
see lenses. Users are also assessed and trained with simulated lenses and known
empty fields. By having thousands of users analyze a survey for short amounts
of time each, it is hoped that a more complete sample of lenses can be
discovered, which can then be fed into lens-finding algorithms to further
improve their performance.

@@

We plan to train a classifier on this data using two different methods. First, we will code 
our own convolutional net in python using theano. Second, we will apply transfer learning 
techniques to train on a convolutional net galaxy morphology classifier, which has graciously 
been made available to us by Ryan Keisler and which achieved 7th place in the 2014 Galaxy 
Zoo Kaggle competition. This classifier runs $96\times96$ images through 3 convolutional 
layers and 2 fully-connected layers and predicts a galaxy to have one of 37 enumerated 
morphologies. We plan to train our classifier on top of the first full-connected layer, which 
has 500 neurons.

@@

We have begun constructing our own convolutional network with two convolutional/max-pool 
layers and two fully-connected layers. Although we have not yet trained on our lensing images,
we can load these images and run them through our initialized classifier.

>>>>

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1}
\begin{document}

@@

Unfortunately, for how useful strong gravitational lenses are, they are also
extremely rare. A next generation optical survey like the Large Synoptic Survey
Telescope can expect to find only ten thousand lenses in the whole sky, while
it will find ten billion galaxies. Currently in astronomy there are only order
hundreds of strong gravitational lenses known, mostly discovered by 'eyeball
squads' of graduate students. The small number means that target criteria must
be somewhat broad in order to maintain a relatively high completeness. Using
reasonable target criteria to find strong lenses such as looking only at
massive galaxies still means that nearly ten million objects will need to be
inspected in the next generation in order to find those ten thousand lenses. A
team of ten graduate students could expect to spend about 14 years looking at
these objects. Computer algorithms are not much better: most current machine
learning algorithms are woefully-underpowered for this task, and generally have
poor completeness or poor purity -- and often poor both. Additionally, some
algorithms are better at finding some types of lenses than others; some
perform well on the brilliant arcs, but poorly on the multiply-imaged objects,
or vice versa. New algorithms need to be developed to find more strong
gravitational lenses, and more strong lenses need to be found to power these
algorithms.

{\sc Space\,Warps}\xspace (Marshall et. al, in prep.) is a citizen-science
initiative designed to overcome these two problems. The program has users
examine images from the Canada-France-Hawaii Telescope Legacy Survey (CFHTLS)
and vote on where they see lenses. Users are also assessed and trained with
simulated lenses and known empty fields. By having thousands of users analyze a
survey for short amounts of time each, it is hoped that a more complete sample
of lenses can be discovered, which can then be fed into lens-finding algorithms
to further improve their performance.

\begin{figure*}[!ht]
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=figs/gallery/0.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=figs/gallery/1.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=figs/gallery/2.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=figs/gallery/3.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
  \caption{Typical Space Warps duds. Insets indicate regions where volunteers typically
  clicked.}
  \label{fig:duds}
\end{figure*}

\begin{figure*}[!ht]
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=figs/gallery/4.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=figs/gallery/5.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=figs/gallery/6.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=figs/gallery/7.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
  \caption{Typical sims. Insets indicate the location of the lens in the
  image.}
  \label{fig:sims}
\end{figure*}

@@

It is clear that we do not have enough data. Luckily, we also know that our
lens objects must obey certain symmetry properties, so it is quite easy to
augment our data. For example, we know that strong lens systems should be
independent of rotations as well as small amounts of stretching and
translation, so our data can be augmented by applying those transformations to
our images.

We plan to train a classifier on this data using two different methods. First,
we will code our own convolutional net in python using theano. Second, we will
apply transfer learning techniques to train on a convolutional net galaxy
morphology classifier, which has graciously been made available to us by Ryan
Keisler and which achieved 7th place in the 2014 Galaxy Zoo Kaggle competition.
This classifier runs $96\times96$ images through 3 convolutional layers and 2
fully-connected layers and predicts a galaxy to have one of 37 enumerated
morphologies. We plan to train our classifier on top of the first
fully-connected layer, which has 500 neurons.

@@

We have created the cutouts of our images from the catalogs.
We have begun constructing our own convolutional network with two
convolutional/max-pool layers and two fully-connected layers. Although we have
not yet trained on our images, we can load these images and run them
through our initialized classifier. Figures~\ref{fig:duds} and
\ref{fig:sims}\xspace show example fields with cutouts inlaid.
